---
title: "Summary statistics of PMLB datasets"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs_sources/")
  })
twitterImg: imgs/summary-plotly.png
---


The plot below is created with [plotly](https://plotly.com/).
Each dot represents a dataset colored based on its associated task (classification vs. regression).
In log scale, the *x* and *y* axis shows the number of observations and features respectively.
Please click on the legend to hide/show the groups of datasets.
Click on each dot to access the dataset's [pandas-profiling](https://pandas-profiling.github.io/pandas-profiling/docs/master/rtd/) report.

```{r setup, include=FALSE}
library(plotly)
library(tidyverse)
library(purrr)
library(htmlwidgets)
# library(htmltools)
library(jsonlite)
library(shiny)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=8.5}
sum_stats <- read_csv('https://raw.githubusercontent.com/EpistasisLab/penn-ml-benchmarks/PMLB2.0/datasets/all_summary_stats.csv') %>% 
  mutate(urls = paste0(
    'https://epistasislab.github.io/penn-ml-benchmarks/profile/',
    dataset,
    '.html'),
    Metadata = paste0('https://github.com/EpistasisLab/penn-ml-benchmarks/tree/PMLB2.0/datasets/', dataset, '/metadata.yaml'))

ply_dat <- sum_stats %>% 
  add_count(problem_type) %>% 
  mutate(Task = paste0(problem_type, ' (', n, ')')) 

ggp <- ply_dat %>% 
  ggplot(aes(x = `#instances`, y = `#features`, color = Task)) +
  geom_point() +
  ggthemes::theme_solarized() +
  # theme_bw() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        # panel.grid.major = element_line(colour = "grey50"),
        panel.background = element_rect(fill = '#FFFFF8'),
        plot.background = element_rect(fill = '#FFFFF8'),
        # legend.background = element_rect(fill = '#FCFCFC')
        ) +
  scale_x_log10(name = 'Number of observations',
                ) +
  scale_y_log10(name = 'Number of features') +
  scale_color_brewer(palette = 'Dark2') +
  NULL
ply <- plotly_build(ggp)


class_dats <- sum_stats %>% 
  filter(problem_type == 'classification') %>% 
  pull(dataset)
fig_lay1 <- gsub(pattern = '<br \\/>Task: classification \\(164\\)', replacement = '', ply$x$data[[1]]$text) %>% 
  gsub(pattern = '#instances:\\s+', '', .) %>% 
  gsub(pattern = '<br \\/>#features:\\s+', ' observations<br />', .) %>% 
  paste0(' features')

ply$x$data[[1]]$text <- paste0('<b>', class_dats, '</b><br /><br />', fig_lay1)

reg_dats <- sum_stats %>% 
  filter(problem_type == 'regression') %>% 
  pull(dataset)
fig_lay2 <- gsub(pattern = '<br \\/>Task: regression \\(122\\)', replacement = '', ply$x$data[[2]]$text) %>% 
  gsub(pattern = '#instances:\\s+', '', .) %>% 
  gsub(pattern = '<br \\/>#features:\\s+', ' observations<br />', .) %>% 
  paste0(' features')

ply$x$data[[2]]$text <- paste0('<b>', reg_dats, '</b><br /><br />', fig_lay2)


ply$elementId <- "PlotlyGraph"

javascript <- HTML(paste("var myPlot = document.getElementById('", ply$elementId, "');
                           myPlot.on('plotly_click', function(data){
                           var urls = ", toJSON(split(ply_dat, ply_dat$Task)), ";
                           window.open(urls[data.points[0].data.name][data.points[0].pointNumber]['urls'],'_blank');
                           });", sep=''))  

ply <- prependContent(ply, onStaticRenderComplete(javascript))
ply
# saveWidget(widget=ply, file=file.path('test.html'), selfcontained = FALSE)
```


Browse, sort and search the complete table of summary statistics below:
```{r echo=FALSE}
my_data <- sum_stats %>% 
  mutate(Metadata = paste0("<a href='", Metadata,"'>", dataset,"/metadata.yaml</a>"),
         dataset = paste0("<a href='", urls,"'>", dataset,"</a>"),
         Imbalance_metric = round(Imbalance_metric, 2),
         `#Classes` = ifelse(
           Endpoint_type == 'continuous', NA,
           `#Classes`
         )) %>% 
  arrange(problem_type, dataset) %>% 
  select(Dataset = dataset, 
         `#Observations` = `#instances`, 
         `#Features` = `#features`, 
         `#Classes`, 
         'Endpoint type' = Endpoint_type, 
         'Imbalanceness' = Imbalance_metric,
         Task = 'problem_type',
         Metadata)

DT::datatable(
  my_data,
   escape = FALSE
#   extensions = 'FixedColumns',
#   options = list(
#   dom = 't',
#   scrollX = TRUE,
#   scrollCollapse = TRUE
# )
)
```



