---
title: "Penn Machine Learning Benchmarks"
# date: "`r Sys.Date()`"
# output:
#   # tufte::tufte_html:
#   html_document:
#     theme: sandston
#     toc: true
#     toc_float: true
#     toc_depth: 2
#     highlight: tango
#     number_sections: true
# knit: (function(inputFile, encoding) {
#     rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs_sources/")
#   })
# twitterImg: imgs/summary-plotly.png
---


This repository contains the code and data for a large collection of curated benchmark datasets for evaluating and comparing supervised machine learning algorithms.
These datasets cover a broad range of applications including binary/multi-class classification and regression problems as well as combinations of categorical, ordinal, and continuous features.


## Summary statistics of PMLB datasets

The plot below is created with [plotly](https://plotly.com/).
Each dot represents a dataset colored based on its associated task (classification vs. regression).
In log scale, the *x* and *y* axis shows the number of observations and features respectively.
Please click on the legend to hide/show the groups of datasets.
Click on each dot to access the dataset's [pandas-profiling](https://pandas-profiling.github.io/pandas-profiling/docs/master/rtd/) report.

```{r setup, include=FALSE}
library(plotly)
library(tidyverse)
library(purrr)
library(htmlwidgets)
# library(htmltools)
library(jsonlite)
library(shiny)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=8.5}
sum_stats <- read_csv('https://raw.githubusercontent.com/EpistasisLab/penn-ml-benchmarks/PMLB2.0/datasets/all_summary_stats.csv') %>% 
  mutate(urls = paste0(
    'https://epistasislab.github.io/penn-ml-benchmarks/profile/',
    dataset,
    '.html'),
    Metadata = paste0('https://github.com/EpistasisLab/penn-ml-benchmarks/tree/PMLB2.0/datasets/', dataset, '/metadata.yaml'))

ply_dat <- sum_stats %>% 
  add_count(problem_type) %>% 
  mutate(Task = paste0(problem_type, ' (', n, ')')) 

ggp <- ply_dat %>% 
  ggplot(aes(x = `#instances`, y = `#features`, color = Task)) +
  geom_point() +
  # ggthemes::theme_solarized() +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        # panel.grid.major = element_line(colour = "grey50"),
        # panel.background = element_rect(fill = '#FFFFF8'),
        # plot.background = element_rect(fill = '#FFFFF8'),
        # legend.background = element_rect(fill = '#FCFCFC')
        ) +
  scale_x_log10(name = 'Number of observations',
                ) +
  scale_y_log10(name = 'Number of features') +
  scale_color_brewer(palette = 'Dark2') +
  NULL
ply <- plotly_build(ggp)


class_dats <- sum_stats %>% 
  filter(problem_type == 'classification') %>% 
  pull(dataset)
fig_lay1 <- gsub(pattern = '<br \\/>Task: classification \\(164\\)', replacement = '', ply$x$data[[1]]$text) %>% 
  gsub(pattern = '#instances:\\s+', '', .) %>% 
  gsub(pattern = '<br \\/>#features:\\s+', ' observations<br />', .) %>% 
  paste0(' features')

ply$x$data[[1]]$text <- paste0('<b>', class_dats, '</b><br /><br />', fig_lay1)

reg_dats <- sum_stats %>% 
  filter(problem_type == 'regression') %>% 
  pull(dataset)
fig_lay2 <- gsub(pattern = '<br \\/>Task: regression \\(122\\)', replacement = '', ply$x$data[[2]]$text) %>% 
  gsub(pattern = '#instances:\\s+', '', .) %>% 
  gsub(pattern = '<br \\/>#features:\\s+', ' observations<br />', .) %>% 
  paste0(' features')

ply$x$data[[2]]$text <- paste0('<b>', reg_dats, '</b><br /><br />', fig_lay2)


ply$elementId <- "PlotlyGraph"

javascript <- HTML(paste("var myPlot = document.getElementById('", ply$elementId, "');
                           myPlot.on('plotly_click', function(data){
                           var urls = ", toJSON(split(ply_dat, ply_dat$Task)), ";
                           window.open(urls[data.points[0].data.name][data.points[0].pointNumber]['urls'],'_blank');
                           });", sep=''))  

ply <- prependContent(ply, onStaticRenderComplete(javascript))
ply
# saveWidget(widget=ply, file=file.path('test.html'), selfcontained = FALSE)
```

Browse, sort, filter and search the complete table of summary statistics below.
Click on the dataset's name to access the dataset’s pandas-profiling report.
Click on the GitHub Octocat <i class="fab fa-github"></i> to access its metadata.
To filter, please type in the box at the bottom of each numeric column in the format `low ... high`.
For example, if you want to see all classification datasets with 80 to 100 observations, type `80 ... 100` at the bottom of the `n_obs` column and select `classification` at the bottom of `Task`.


```{r echo=FALSE}
my_data <- sum_stats %>% 
  mutate(Metadata = paste0("<a href='", Metadata,"'>", 
                           as.character(icon("github", lib = "font-awesome")),"</a>"),
         dataset = paste0("<a href='", urls,"'>", dataset,"</a>"),
         # Dataset = paste(dataset, Metadata),
         `Imbalance` = round(Imbalance_metric, 2),
         n_classes = ifelse(Endpoint_type == 'continuous', NA, `#Classes`),
         Task = as.factor(problem_type),
         Endpoint = as.factor(Endpoint_type)) %>% 
  arrange(problem_type, dataset) %>% 
  select(Dataset = dataset, n_observations = `#instances`, n_features = `#features`, 
         n_classes, Endpoint, `Imbalance`, Task, Metadata)

DT::datatable(
  my_data,
   escape = FALSE,
  extensions = 'Buttons', options = list(
    dom = 'Blfrtip',
    buttons = c('csv')
  ),
  filter = list(position = 'bottom', clear = FALSE),
  rownames= FALSE,
#   extensions = 'FixedColumns',
#   options = list(
#   dom = 't',
#   scrollX = TRUE,
#   scrollCollapse = TRUE
# )
)
```

## Dataset format

All datasets are stored in a common format:

* First row is the column names
* Each following row corresponds to one observation of the data
* The dependent variable/endpoint column is named `target`
* All columns are tab (`\t`) separated
* All files are compressed with `gzip` to conserve space


## Citing PMLB

If you use PMLB in a scientific publication, please consider citing the following paper:

Randal S. Olson, William La Cava, Patryk Orzechowski, Ryan J. Urbanowicz, and Jason H. Moore (2017). [PMLB: a large benchmark suite for machine learning evaluation and comparison](https://biodatamining.biomedcentral.com/articles/10.1186/s13040-017-0154-4). *BioData Mining* **10**, page 36.

BibTeX entry:

```bibtex
@article{Olson2017PMLB,
    author="Olson, Randal S. and La Cava, William and Orzechowski, Patryk and Urbanowicz, Ryan J. and Moore, Jason H.",
    title="PMLB: a large benchmark suite for machine learning evaluation and comparison",
    journal="BioData Mining",
    year="2017",
    month="Dec",
    day="11",
    volume="10",
    number="1",
    pages="36",
    issn="1756-0381",
    doi="10.1186/s13040-017-0154-4",
    url="https://doi.org/10.1186/s13040-017-0154-4"
}
```

## Support for PMLB

PMLB was developed in the [Computational Genetics Lab](http://epistasis.org/) at the [University of Pennsylvania](https://www.upenn.edu/) with funding from the [NIH](http://www.nih.gov/) under grant R01 AI117694.
We are grateful for the support of the NIH and the University of Pennsylvania during the development of this project.

