---
title: "R interface"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    css: custom.css
---

pmlbr :package: provides the following user-facing functions and variables in the R environment.
For installation instructions or a high-level overview of how to use PMLB, please see [Using pmlbr](./using-r.html).

## Functions

### `fetch_data` {#docstring-output}

*function* `pmlb.fetch_data(dataset_name, return_X_y=False, local_cache_dir=None, dropna=True)`

Download a dataset from PMLB, (optionally) store it locally, and return the dataset.

You must be connected to the internet if you are fetching a dataset that is not cached locally.

**Parameters:**

- **dataset_name : str**
  Name of the dataset to fetch from PMLB. A list of available datasets can be found in the documentation.
- **return_X_y : bool, default=FALSE**
  Specify the format of the data returned to the user. If `FALSE`, the data are returned as a dataframe of size `(n_samples, n_features+1)`, where the last column is `target` - the true target value of the data point. If `TRUE`, the data are returned as a list of length 2, where the first element is a dataframe of size `(n_samples, n_features)` corresponding to the feature matrix (usually denoted `X`), and the second element is a numeric vector of length `n_samples` corresponding to the list of target values (usually denoted `y`), respectively.
- **local_cache_dir : str, default=None**
  The directory on your local machine in which to store the data files, defaults to NA, indicating cache will not be used.
- **dropna : bool, default=TRUE**
  If `TRUE`, `NA`s will be automatically dropped from the exported dataset.

**Returns:**

- **dataset : dataframe**
  If `return_X_y == FALSE`, fetched dataset as a dataframe of size `(n_samples, n_features+1)`, where the last column is `target`. If `return_X_y == TRUE`, a list of length 2 containing the feature matrix `X` and the target array `y`, respectively.

**Example:**

```{r}
library(pmlbr)

# Download features and labels for penguins dataset in single data frame
penguins <- fetch_data('penguins')
str(penguins)
```

## Data objects

These objects contain details on datasets currently available in PMLB. 

### `dataset_names` {#docstring-output}

A list of all datasets included in PMLB.

The contents of this vector are the union of `classification_dataset_names` and `regression_dataset_names`.

**Example:**

```{r}
data(pmlb)
head(dataset_names, 9)
```

### `classification_dataset_names` {#docstring-output}

A list of all *classification* datasets included in PMLB.

Classification datasets are datasets where the target values (dependent variable/outcome) are discrete.

**Example:**

```{r}
head(classification_dataset_names, 9)
```

### `regression_dataset_names` {#docstring-output}

A list of all *regression* datasets included in PMLB.

regression datasets are datasets where the target values (dependent variable/outcome) are continuous.

**Example:**

```{r}
head(regression_dataset_names, 9)
```

### `summary_stats` {#docstring-output}

Summary statistics for the all datasets.

A data frame with 10 variables:

- *dataset*: Dataset name
- *n_instances*: Number of data observations (equal to number of rows)
- *n_features*: Total number of features (number of columns - 1)
- *n_binary_features*: Number of binary features
- *n_categorical_features*: Number of categorical features
- *n_continuous_features*: Number of continuous features
- *n_classes*: Number of classes in target variable
- *endpoint_type*: Value type of endpoint/target (can be binary, categorical or continuous)
- *imbalance*: Imbalance measure, where zero means that the dataset is perfectly balanced and the higher the value, the more imbalanced the dataset
- *task*: Type of problem/task. Can be classification or regression.

**Example:**

```{r}
head(summary_stats)
```
