---
title: "Using pmlbr"
output:
  html_document:
    toc: false
---

## Install pmlbr

The [**pmlbr** package](https://github.com/EpistasisLab/pmlb) :package: provides the following user-facing functions and variables in the R environment.

The package works for any recent version of R.

You can install the released version of **pmlbr** from CRAN with:
```{r, eval=FALSE}
install.packages('pmlbr')
```

Or the development version from GitHub with remotes:

```{r, eval=FALSE}
# install.packages('remotes') # uncomment to install remotes
library(remotes)
remotes::install_github("EpistasisLab/pmlbr")
```

## Usage

The core function of this package is `fetch_data` that allows us to download data from the PMLB repository.
For example:
``` {r}
library(pmlbr)

# Download features and labels for penguins dataset in single data frame
penguins <- fetch_data('penguins')
str(penguins)

# Download features and labels for penguins dataset in separate data structures
penguins <- fetch_data('penguins', return_X_y=TRUE)
head(penguins$x) # data frame
head(penguins$y) # vector
```

Let's check other available datasets and their summary statistics:

``` {r}
# Dataset names
head(classification_dataset_names, 9)
head(regression_dataset_names, 9)

# Dataset summaries
head(summary_stats)
```

Selecting a subset of datasets that satisfy certain conditions is straight forward with `dplyr`.
For example, if we need datasets with fewer than 100 observations for a classification task: 
```{r warning=FALSE, message=FALSE}
library(dplyr)
summary_stats %>% 
  filter(n_instances < 100, task == 'classification') %>% 
  pull(dataset)
```


### Dataset format

All data sets are stored in a common format:

* First row is the column names
* Each following row corresponds to an individual observation
* The target column is named `target`
* All columns are tab (`\t`) separated
* All files are compressed with `gzip` to conserve space

This R library includes summaries of the classification and regression data sets but does **not**
store any of the PMLB data sets.  The data sets can be downloaded using the `fetch_data` function which
is similar to the corresponding PMLB python function.  

Further info:
``` {r}
?fetch_data
?summary_stats
```


### Citing

If you use PMLB in a scientific publication, please consider citing the following paper:

Randal S. Olson, William La Cava, Patryk Orzechowski, Ryan J. Urbanowicz, and Jason H. Moore (2017).
[PMLB: a large benchmark suite for machine learning evaluation and comparison](https://biodatamining.biomedcentral.com/articles/10.1186/s13040-017-0154-4).
BioData Mining 10, page 36.


## Roadmap

* Add tests
* Submit library to CRAN


## Contributing

Pull requests are welcome.  For major changes, please open an issue first to discuss what you would like to change.

Integration of other data repositories are particularly welcome.
