---
title: "Using pmlbr"
output:
  html_document:
    toc: false
---

## Install

The [**pmlbr** package](https://github.com/EpistasisLab/pmlb) provides the following user-facing functions and variables in the R environment.

The package works for any recent version of R.

You can install the released version of **pmlbr** from CRAN with:
```{r, eval=FALSE}
install.packages('pmlbr')
```

Or the development version from GitHub with remotes:

```{r, eval=FALSE}
# install.packages('remotes') # uncomment to install remotes
library(remotes)
remotes::install_github("EpistasisLab/pmlbr")
```

## Usage

The core function of this package is `fetch_data` that allows us to download data from the PMLB repository.
For example:
``` {r}
library(pmlbr)

# Download features and labels for penguins dataset in single data frame
penguins <- fetch_data('penguins')
str(penguins)

# Download features and labels for penguins dataset in separate data structures
penguins <- fetch_data('penguins', return_X_y=TRUE)
head(penguins$x) # feature data frame
head(penguins$y) # target vector
```

Let's check other available datasets and their summary statistics:

``` {r}
# Dataset names
head(classification_dataset_names, 9)
head(regression_dataset_names, 9)

# Dataset summaries
head(summary_stats)
```

Selecting a subset of datasets that satisfy certain conditions is straight forward with `dplyr`.
For example, if we need datasets with fewer than 100 observations for a classification task: 
```{r warning=FALSE, message=FALSE}
library(dplyr)
summary_stats %>% 
  filter(n_instances < 100, task == 'classification') %>% 
  pull(dataset)
```

Further info:
``` {r}
?fetch_data
?summary_stats
```
